---
title: "ENVS 193 Final"
author: "Michael Weigand"
date: "2026-03-17"
output: html_document
---

[Link to my GitHub Repository](https://github.com/michiwei2002-bit/ENVS-193DS_winter-2026_final)


library(janitor)



Problem 1

a. Transparent statistical methods
In part 1, they used a simple linear regression. In part 2, they used a Kruskal-Wallis test.

b. Figure needed
They should make a boxplot. The x-axis should be the water year classification (wet, above normal, below normal, dry, critical drought), and the y-axis should be the flooded wetland area (m²). The visual components would include boxes representing the interquartile range for each water year category, with a thick center line explicitly showing the median flooded area, and whiskers demonstrating the overall spread of the data.

c. Suggestions for rewriting
Precipitation did not significantly predict flooded wetland area in the San Joaquin River Delta (Linear regression: test statistic = [value], df = [value], p = 0.11). Furthermore, the median flooded wetland area did not significantly differ across the five water year classifications (Kruskal-Wallis: test statistic = [value], df = [value], p = 0.12).

d. Interpretation
One additional variable that could influence the flooded area is "Active Water Management Status," which is a categorical variable (e.g., actively managed vs. unmanaged). This would influence wetland flooding because managers frequently pump water into these habitats artificially to support breeding waterfowl, overriding natural precipitation levels or drought conditions.


Problem 2

```{r}
kelp <- read_csv(here("Data", "Annual_Kelp_All_Years_20250903.csv"))
```


a.
```{r}
kelp_clean <- kelp %>%
  # 1. Standardize column names to snake_case so they are easy to type
  clean_names() %>% 
  
  # 2. Keep only the data for the three specific sites we care about
  filter(site %in% c("IVEE", "NAPL", "CARP")) %>% 
  
  # 3. Remove this specific anomalous data point
  filter(!(date == "2000-09-22" & transect == "6" & quad == "40")) %>% 
  
  # 4. Create a new column with the full, readable names of the sites
  mutate(site_name = case_when(
    site == "NAPL" ~ "Naples",
    site == "IVEE" ~ "Isla Vista",
    site == "CARP" ~ "Carpinteria"
  )) %>% 
  
  # 5. Convert that new column to a factor and explicitly set the order for the plot legend
  mutate(site_name = as_factor(site_name),
         site_name = fct_relevel(site_name, "Naples", "Isla Vista", "Carpinteria")) %>% 
  
  # 6. Group the data by site and year so we can calculate annual averages per site
  group_by(site_name, year) %>% 
  
  # 7. Calculate the mean number of fronds for those groups
  summarize(mean_fronds = mean(fronds)) %>% 
  
  # 8. Ungroup the data to prevent accidental grouped math later on
  ungroup()

# Display the required outputs
slice_sample(kelp_clean, n = 5)
str(kelp_clean)
```

